A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.

### 2019-12-09 paper.Rmd
final changes before review by Karthik

To share data analysis data, there should be some minimal set of requirements. For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)). No matter where data is submitted, there should ideally be a canonical data repo in one long term archive that links to others with a DOI.

@@

Another rich format is XML, the e**X**tensible **M**arkup **L**anguage. This was an early iteration on the idea of a plain text format that was both human and machine readable. It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable. XML is used is in EML, the Ecological Metadata Language. XML differs to JSON in that JSON is shorter, and quicker and easier to read and write. JSON can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser. JSON has key-value pairs of information, whereas XML represents data in a tree structure. XML and JSON can be converted into each others respective formats, for example with the `json` or `xml2` R packages.

@@

If the dataset you are shipping has a research application, the most relevant home for it would be the research data repositories **zenodo** or **dryad**. Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets. The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases. Zenodo is able to accommodate larger file sizes upon request. Dryad will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, but requires a data publishing fee.

@@

1. **Decide whether publishing is appropriate for your data** 
It is critical to publish your data if they are the basis for a peer-reviewed research publication. To be broadly useful, your data must be deposited in a permanent archive, to ensure that it does not disappear from a ephemeral location such as your university website. It shoudl also contain useful metadata so that someone outside the project team can understand what the data mean. However, this level of effort is not always critical for all data science efforts. For small-scale projects such as ones where one might generate simulated datasets, this level of curation is highly unnecessary. Here making data available in a transient location like GitHub or a website is sufficient.

>>>>

To share data analysis data, there should be some minimal set of requirements. For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)). No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others with a DOI.

@@

Another rich format is XML, the e**X**tensible **M**arkup **L**anguage. This was an early iteration on the idea of a plain text format that was both human and machine readable. It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable as JSON. JSON is shorter, and quicker and easier to read and write than XML, and can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser. XML and JSON can be converted into each others respective formats, for example with the `json` or `xml2` R packages. XML is still used as metadata for data storage, for example in EML, the Ecological Metadata Language [@EML-about].

@@

If the dataset you are shipping has a research application, the most relevant home would be the research data repositories **zenodo** or **dryad**. Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets. The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases. Zenodo is able to accommodate larger file sizes upon request. The Dryad Digital Repository (@dryad) will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, in exchange for a data publishing fee.

@@

1. **Decide whether publishing is appropriate for your data** 
It is critical to publish your data if they are the basis for a peer-reviewed research publication. To be broadly useful, your data must be deposited in a permanent archive, to ensure that it does not disappear from a ephemeral location such as your university website. It should also contain useful metadata so that someone outside the project team can understand what the data mean. However, this level of effort is not always critical for all data science efforts. For small-scale projects such as ones where one might generate simulated datasets, this level of curation is highly unnecessary. Here making data available in a transient location like GitHub or a website is sufficient.
